---
title: "caclec_5-8-2019_comments"
author: "grady"
date: "May 8, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!--
# Intro Thoughts

Interested to see what people think.  In order to not influence the conversation, I will reserve them for later.

# Other's Thoughts

```{r, eval=F}
set.seed(347)
us <- c("Nuole", "Charla", "Ekrem", "Cila", "Alice")
sample(us)

```


# Ekrem

- many little comments scattered across document.
- much improved

### General

- don't know if framing it by going to other itnerventions is good.  good to talk about other intervention, and how these mechanisms can augmnet/complement.  But not comparison/competition with other interventions.
    - (1) when we talk about interventions, a number of different interventions.  peacekeeping/UN, mediation/diplomacy, etc...these are all aimed at diff things.  information problems, security guarantees, increasing costs of reneging, etc... This is a super heterogeneous group. Not clear which thing my paper is about and should be compared with.
    - (2)
- could set up with trust problems.  That's the more general thing.  THe connection from trust to formal/rationalist stuff is easy to make.  The connection between trust and my psych stuff is super easy to make.  Mechanisms of increasing trust and reducing prejudice.
    - information/comittment problems, tie to trust, tie to my stuff.  Deeper mechanism, being talked about in ps.
- The theory part (the current bottom of the paper/theory section) should be my pitch.  The connections that will be built through a way to increase trust.  NOt about itnerventions alone -- they are a subcategory of trust.  Increasing trust can help in fundamental ways.  Emphasize the complementarity of international itnerventions and bottom-up stuff.  Make them complements and not substitutes.
- (1) why trust is such an important issue in these conflicts.  two ways in which intervention is helping.  (1) increasing trust directly, helpful for lots of reasons.  (2) cleaning the pipes argument -- we would be able to have contact and learn and update rationally, but prejudices make the pipes clogged.  Information comes through biased.
    - then intervention is a subcategory.  If we think of contact & trust this way, we can think of itnerventions and other stuff this way.

# Cila

- Likes intervention framework.  Comments are mostly about clarifying it.
- After hearing ekrem, maybe trust framework is a good idea.
    - even with trust framework in beginning, more nuanced argument about international itnervention.  Maybe over-generalizing all interventions as top-down -- overly negative tone about international intervention.
- more transition between bottom-up peacebuilding interventions and THIS intervention.  Not clear if THIS intervention is a part of the bottom-up approach.  And if there are still problems for bottom-up peacebuilding, what are they? How does my approach solve those?
- **Biggest comment**: calling it a psychological intervention.  WHen explaining the experiment, they build a lot of infrastructure and stuff.  So it seems like an infrastructure project.  ANd people who did not get intervention had positive impact -- wouldn't that be because of the infrastructure and increasing quality of life?
    - **me**: superordinate goals having positive externalities.
- and not clear why superordinate goals make contact more successful.  Not well explained.
- A hypothesis about increased contact.  Why?  Isn't that the treatment?

# Charla

- improved even from the Aleks lab version of the intro.
- less of an issue with the top-down negativity when reading, BUT definitely framed as top-down is bad.  Will be a bigger issue for discussion.
- Introduction is too long.  Probably needs to shorten.  But good to spell out the contributions, because there are a lot of contirbutions of this paper.  ANd this section makes them clear.  ANd there's a clear connection between those contributions and the paper's discussion/theory.
- Lots of hypotheses: breaking them apart into multiple papers?  WHat doing with them?

- Talk a lot about bargaining problems -- information, comittment, reputation, etc...Disconnect knowing how I set up the experiment.  Big tone shift between front-end and theory part.  A bit jarring.  Theory is less developed/edited than the intro.  Often "statement 1, statement 2" but no logical connection between that.
- Primary audience is probably not people who are well versed in comittment/information/reputation problems.  So need to make those connections clear.

# Nuole

- Agree with Charla about too long introduction.  And my contributions don't come early enough, but it is excellent.  A particular section reads too much like a lit review, though.
- Likes intervention framing, but the project is not really about international intervention.
- didn't think too negative, but was thinking about complentarity with top-down and bottom-up from beginning.
- For framing: two groups (A and B), have some info/comittment problem which causes conflict.  To resolve problem, mediation and international intervention are 3rd party ways to solve comittment and information problems.  _Complement to that_ is bottom-up.  Conflicts that ar enot big enogh for international community to mediate.  Lots of "lower level" conflicts which could be resolved with bottom-up approaches.  A way to solve it from bottom-up is intergroup contact/contact hypothesis/prejudice reduction.
    - both deal with information and comittment problems to resolve conflict.
    - Saying "here's one way we solved these bargaining problems, let me show you another way".  ANd let me tell you how it can be better/complementary.
- Want more of the trust/psych stuff up front.  Too much stuff about intervention up front so everyone thinks its about internaitonal itnervention.  Need to bring up more bargaining problems/trust earlier.
- But generally likes the setup/layout.

- Really likes introduction with the villages.  
    - Charla agrees.  For that reason, try not to change first 3 paragraphs.  Move 4th paragraph out.  Summarize more about the study and then get into contributions.


# Discussion

- feels a disconnect between the research and the intro/framing at times.
- WHen we think about top-down interventions, we usually think about with _organized sides_.  Shuttle diplomacy: I am talking with rebel leader, I am talking with government.  But this is communal violence.
    - problematic with _current_ rhetoric.  But if change it a little bit to complementarity with intervention, this difference is a strength.  **How can we even do this intervention/mediation when there are no leaders??**  It is really not obvious what a top-down intervention _even is_.  We should learn more on communal trust building. [Ekrem].
    - Contact theory: works cooperate, equal status, shared goals, support of elites.  What is the role of elites?  Are they formal leaders?  Who are they/what is their role?  Without elites, same effect?  Leading people to do things or what?
        - I actually do not think this is well theorized.  But about social norms.  Charla says address this in maybe 1-2 sentences but not much.
        - Link ROle of elites with Mediation training component of program.
        - Nuole says: what is it if people change attitudes without elites?  A revolution?

- Ekrem: when talking about making the paper about TRUST instead of intervention, not a big change.  Just moving around a few things.
    - mediation always starts with: "what would be the way out for these groups by themselves?  If we add a mediator, how cna this improve the current situation?".  Good roadmap for me to follow perhaps.  Lay out comittment problem/information problem -- weak states (like Nigeria) that cannot solve the problem.  SO how can the communities solve this themselves?  Reputation/comittment problems. 
    - First lay that out.  Then discuss my contributions about that udnerlying stuff.  If we start with intervention, it is about intervention.  It can be at a more fundamentla level.  "We alerady have a way to think about solving these problems.  Here is a thing that (1) makes those interventions work better.  And (2) where the interventions are too costly/not feasible, we can still do stuff."
        - top-down interventions should work better with these mchasnisms.  We are cleaning the pipes so ANYONE bargaining will do it better, process information less biasedly.
        - Nuole: Get 2.1 in the intro.  It connects the paper to the trust/bargaining problems.  2.1 is the bigger stuff about bargaining, the interventions are the example of bargaining.
        - Ekrem: and just use the word trust instead of reputations.

- Ekrem: 3 works in his comments.  
    - Andew Kydd's book (maybe 2015) "Trust and Mistrust in Internaitonal Relations" and WOrld Politics article in 1997.  Book is trust between SOviets and US taking small steps to icnrease trust between each other.
    - JCR and Economics article.  Directly looking at civil wars.  Both have models where individual people within different groups are randomly interacting in some way.  Overall trust across these groups determines the outcomes of these individual pairings.  What each individual does has an effect ont he overall trust these groups have for each other.  Trust FOR your group is a public good.  Look at these trust relatinoships and how these can lead to vicious/virtuous cycles for community relationships.
    - Those 3 articles can complement fearon/powell for framing the initial stuff.  
        - The mediation stuff is after that.  Second-level, an example of these other things.


# My Thoughts

- some paragraphs (or large parts of paragraphs) I can cut.
    - "International intervention through military intervention and mediation might be conducive to preventing different types of violence than bottom-up peacebuilding programs." That paragraph is not good.  I think there is something there but...maybe revised and relegated to discussion section, not intro.
    - "Local turn" paragraph is good but has some bad parts.  Sort of sets the paper as part of the bottom-up vs top-down debate in peacekeeping, which is not really what I want.

- The intro "this paper contributes to the literature on" section is weak.
    - Fulani paragaph is _real weak._
    - **people said this is one of the substantive reasons we shuld care about.**

- the top-down vs bottom-up/grassroots comparison.  Something is overly antagonistic about it.  Was reading a lot of Autesserre, who is super critical of UN peacekeeping's top-down approach.  But I don't think that's the comparison I want to make.  The paper is not comparing top-down and bottom-up programs, and I am not really trying to say that top-down is bad and bottom-up is good.
    - **language is "substitutey", should be "complementy"

- I spend too much Theory time in the "intergroup bargaining and why intervention sometimes fails" mode.  I need to get to the "how prejudice complicates intergroup bargaining" and "how psychological reconciliation can reduce prejudice and assist intergroup bargaining."
    - Also the words I am using are not quite right: top-down, bottom-up, psychological reconciliation, etc...
    - **move 2.1 to intro so real theory comes earlier**

- One of my concerns is that the traditional peacebuilding literature focuses on civil wars and this project is not civil conflict.  Does not involve state actor.
    - **If framing this paper more about trust, then civil wars and intervention are one example of trust issues and MY STUFF is another example.  Not opposed, an example of the same problem.**
    - I am not differentiating between my program and some other type of program.  Not about types of itnerventions, so dont set it up as one type vs another type.
    - If framing at more general level and then moving to interventions, can talk about interventions a lot less.  BUndle.  Don't need to be so nuanced.  ALl of these interventions help through certain channels.  my intervention helps people directly but also indirectly.  Can be more succint with the interventions.
    - Intro is short story from Nigeria, stuff about trust/bargaining issues, then get to summary of paper (expeirment and my contributions).  Intervention stuff is a contribution, not main point.

- Another concern is placing our intervention in the peacebuilding literature but keeping the focus on the psychological nature of the intervention.  

- Another concern is the conflation of peacebuilding/peacekeeping/peacemaking.
-->

# Summary of comments

The most useful positive comments were (1) how this paper brought together theories of conflict from Econ/IR and theories of prejudice reduction from Psych/American Politics.  People thought this was huge and a big contribution.  The psych/econ ideas clearly work together well but very few papers bring them together like this.  And (2) the first three paragraphs of the intro, using the story of farmers protecting pastoralists, really brought people into the paper.

The most useful "negative/constructive" comments were about moving the framing one more step up.  Specifically, the framing should be about solving the trust problems of intergroup bargaining, not about types of international interventions.  Focusing on trust problems is great for us.  It fits much better into most PS literature about intergroup conflict as a trust problem.  And it fits much better with the theory section bringing ideas from psychology to help solve trust problems.  Trust problems are the bigger phenomenon that international interventions seek to address, so that stuff isn't totally off topic, but should not be the focus.

I think this was most the most useful set of comments overall because it solves most of the framing problems I'd been thinking about after taking a day off from the paper: (i) The awkwardness of talking about international interventions (in the intro) before talking about intergroup bargaining (was first section of theory); (ii) Framing with a literature on intervention that concerns organized groups and applying it to communities (solved if we frame it around trust problems that the communities have, and use interventions as examples of these trust problems); (iii) The semi antagonistic "top-down" vs "bottom-up" comparison that we don't want to make and that our theory is not about; (iv) the focus being on other interventions instead of the theory of prejudice reduction helping solve bargaining problems that lead to conflict.

All around just a huge win to make our pitch about the bigger "trust" phenomenon.  Everyone's main comments were some flavor of this, and after discussion the group had consensus on "trust" framing.  Which is a great framing for us!

## Set 1 -- Primary

- Make this less about intervention/peacekeeping and more about the trust/bargaining problems that cause violence.  And how prejudice reduction alleviates trust/bargaining problems (our major contribution).  Not a big change, just moving things around and talking less about interventions.
    - Focus on trust/bargaining problems also makes intergroup trust the center of the analysis, with security/violence-reduction like a behavioral implication/downstream outcome.
    - peacekeeping/building frame is narrower than this needs to be; trust issues are a huge issue in lots of aspects of political science.

- Like the two ways reducing prejudice decreases bargaining problems: (1) increasing trust directly, and (2) "cleaning the pipes" -- prejudice "clogs pipes" and biases information processing, which decreases trust by changing how we interpret the behavior of others.  Reducing prejudice these two ways helps _all_ bargaining.  That is huge.

- Don't want to set up paper as a "top-down" vs "bottom-up" study because it's not.  That's not the theory, not the design, not the point.  We are not comparing two types of interventions.  The language right now is too "substitute" and needs to be more "complement".
    - Also, both "top-down" and "bottom-up" deal with information and commitment problems to resolve conflict.  Solving information and commitment problems is the point.
    - Contrast with top-down interventions is weird because those are targeted to organized states/groups, where a state leader and a rebel leader bargain.  That is not the case here.
    - Strategies of removing prejudice help "top-down" interventions, as well, since less biased information processing helps all bargaining.

- Focus on complementarity of "improve bargaining through prejudice reduction" with other ways of improving bargaining (mediation, punishing defection, etc...).  A way this type of "prejudice reduction" intervention complements other ways to improve bargaining: lots of conflicts are not big enough for international community to mediate and/or lack unified groups where 1 leader can speak for the group.  These "lower level" conflicts can be resolved with bottom-up approaches, whereas it is hard to think of a "top-down" approach to lower-level conflicts.  How can we even do intervention/mediation without unified groups?  A way to solve it from bottom-up is intergroup contact/contact hypothesis/prejudice reduction and building trust.
    - If framing this paper more about trust, then civil wars and international intervention (most lit cited in intro) are one example of trust issues, and Nigeria stuff is another example.  Not a problem that most of the international intervention is about civil wars!  Civil wars and the intercommunal conflict are an example of the same larger problem.

- Some very useful papers about intergroup trust and intergroup bargaining.  Those papers can complement fearon/powell about "intergroup conflict as a bargaining problem" for framing the initial stuff.  The papers have models where individual people within different groups are randomly interacting.  Overall trust between these groups determines the outcomes of these individual pairings.  What each individual does has an effect on the overall trust these groups have for each other.  Trust FOR your group is a public good.  Look at these trust relationships and how these can lead to vicious/virtuous cycles for community relationships.
    - In our paper, the intervention/mediation stuff should come after this bigger trust stuff.  Intervention/mediation is second-level, an example of these bigger trust problems.

- **To implement**: move up section 2.1 to the intro, cut some of the intro about interventions, and be more general/bundled with motivating literature (less specific = less space = more space for the good theory stuff).  This also has the benefit of getting the contribution earlier in the intro, and keeping the theory focused on the prejudice reduction and not reviewing other interventions.  It gets the theory about prejudice earlier, too.
    - Maybe: Lay out commitment problem/information problem, maybe how weak states (like Nigeria) cannot solve the problem the way it is typically solved (mediation/intervention).  So how can the communities solve this themselves?  Reputation to overcome commitment problems. Then discuss contributions about that underlying stuff. "We already have a way to think about solving these problems (reputation/trust, mediation, intervention).  Here is a thing that (1) makes those things work better.  And (2) where those things are too costly/not feasible, we can still do this anti-prejudice stuff."


## Set 2 - Secondary

- People love how this takes two literatures that do not talk to each other (Bargaining from Econ/International relations and prejudice reduction from Psych/American politics) and shows how much we can learn from both.

- Intro is too long.  Definitely keep first three paragraphs.  Keep discussion of contributions at end of intro.  Streamline sections between first three paragraphs and description of study by talking less about international intervention/mediation.
    - Intro should be short story from Nigeria, stuff about trust/bargaining issues, summary of paper (theory, experiment, results, and contributions).  Things about international intervention stuff is a nice contribution, but not main point.

- 2/5 of the group perceived this as overly negative about top-down interventions, which takes away from the paper, is unnecessary for the paper's argument, is not what we believe, and will make enemies.

- More about how the material improvements in the program (boreholes, buildings) translates to the psychological stuff in the theory (this maybe goes into the "intervention design" section that comes before the "research design" section).  
    - Maybe talk about "spillover" as positive externalities from contact & superordinate goals.  10 people from each group get together and everyone gets water (superordinate goal achieved).  The water could make everyone less prejudiced, especially if they know the other side helped bring it.  That is good for a program trying to reduce prejudice and produce positive spillovers.
        - also not clear why superordinate goals make contact more successful.  And the role of elites/leaders in contact is not clear.

## Set 3 - Writing

- Theory is less developed/edited than the intro.  Often "statement 1, statement 2" but no logical connection between that.  When finishing the intro, remember to connect the statements.

- A primary audience for this paper will be people who are not well versed in commitment/information/reputation problems.  So need to make those connections _very_ clear and clearly tied to theory/intervention.
